Text Is All You Need

### 2023-05-11 docs/hes_final/presentation.qmd
Removed some redundant slides

## Three approaches to extracting conflict signals from text embeddings

- The first approach used supervised dimensionality reduction and a feed-forward neural network conditioned with country context characteristics.
- In the second approach, we used a compact dataset of monthly embedding averages per country per month and applied statistical models such as K-Nearest Neighbors and XGBoost.
- In the third approach we stacked multiple article-level embeddings and applied transformer-based models to extract a signal exclusively from the text corpus.

@@

## Introducing our 'spike' conflict metric

- We derived a metric designed to highlight relative “spikes” in political violence. Calculating whether a spike occurred in a given month is a two-step process. 
- First, we calculated the trend in violence for a given month, defined by the slope of OLS regression of the number of ACLED-documented fatalities in target month & preceding 2 months.
- Next, we calculated whether that slope was in the 75th percentile of such slopes in the target country for the preceding 2 years.

:::{.notes}
This trend allowed us to estimate whether violence was increasing or decreasing in the target country at any time.
:::
---

@@

- Our 'spike' binary metric is sensitive to small increases in political violence in peaceful countries/periods, while also filtering out small fluctuations in long-standing conflicts.

@@

- Models, not oracles 
- Not a substitute for careful deliberation and dialogue
- At their best, these models can provide an early warning for conflict, and if the right actions are taken, then they can be made to be wrong

>>>>

@@

## The 'Spike' conflict metric

- We derived a metric to highlight relative “spikes” in political violence.
- Two step calculation
  + First, we calculated the trend in violence for a given month
    - Defined by the slope of OLS regression of the number of ACLED-documented fatalities in target month & preceding 2 months. 
  + Next, we calculated whether that slope was in the 75th percentile of such slopes in the target country for the preceding 2 years.

:::{.notes}
This trend allowed us to estimate whether violence was increasing or decreasing in the target country at any time.
:::

### Spike Metric Cont...

@@

- Sensitive to small increases in political violence in peaceful countries/periods, while also filtering out small fluctuations in long-standing conflicts.

@@

## Mitigating Risk
- Models, not oracles 
- Not a substitute for careful deliberation and dialogue
- At their best, these models can provide an early warning for conflict, and if the right actions are taken, then they can be made to be wrong
